---
title: "Comparing pharma  & clinics"
author: "Josh Herbeck"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r setup, message=FALSE}

library(deSolve)
library(tidyverse)
library(EpiModel)
library(survival)
library(EasyABC)
```

### Model setup  

We are modeling a closed population SI deterministic compartmental model meant to simulate infections over a course of a single year-long trial. We are not modeling infections from I to S but rather only infections from the outside (non-trial) population, with the infection rate based on the population prevalence `prev` (of viremic individuals), the exposure rate (serodiscordant sexual contacts per time) `c`, and the transmission rate (per contact) `p`.  

`beta` = transmission rate (per contact)   
`c` = exposure rate (serodiscordant sexual contacts per time)  
`prev` = prevalence  (prevalence of viremic individuals)  
`lambda = beta * c * prev`  
`risk` = risk multiplier  

The risk multiplier is an amalgam of increases in transmission risk that could be due to: 1) increased per contact transmission risk; 2) increased exposure rate (number of contacts); or 3) increased prevalence of HIV viremia in partners. Individual risk of infection can vary for these separately or in combination.  

Basic functions:  

`dS/dt = -lambda*S`   
`dI/dt = lambda*S`  

Spharma = susceptible pharma  
Ipharma = infected pharma  
Sclinic = susceptible clinic  
Iclinic = infected clinic  

### Calibration, qualitative first step

We can do an initial qualitative calibration to incidence in sub-Saharan Africa using an initial set of transmission parameters borrowing from Alain Vandormael (2018):    

     "We used realistic parameter values for the SIR model, based on earlier HIV studies that have been undertaken in the sub-Saharan Africa context. To this extent, we varied `c` within the range of 50 to 120 sexual acts per year based on data collected from serodiscordant couples across eastern and southern African sites. Previous research has shown considerable heterogeneity in the probability of HIV transmission per sexual contact, largely due to factors associated with the viral load level, genital ulcer disease, stage of HIV progression, condom use, circumcision and use of ART. Following a systematic review of this topic by Boily et al., we selected values for `beta` within the range of 0.003–0.008. ... Here, we chose values for `v` within the range of 0.15–0.35, which are slightly conservative, but supported by population-based estimates from the sub-Saharan African context."

`c` varies from 50 to 120 per year   
`beta` varies from 0.003 to 0.008  
`prev`, which here is population prevalence of unsuppressed VL, varies from 0.15 to 0.35  

### Model function; ODEs  

Right now this includes just two trial populations, one from the clinic and one from the pharmacy. The pharmacy pop has higher risk of infection (defined by the "risk" parameter).  
  
```{r ODEs}

si_ode <- function(times, init, param){
  with(as.list(c(init, param)), {
    
    # Flows
    # the number of people moving from S to I at each time step
    
    #Susceptible, Infected, high, 
    SIpharma.flow <- risk*lambda*Spharma
    SIclinic.flow <- lambda*Sclinic
    
    # ODEs
    # placebo; heterogeneous risk
    dSpharma <- -SIpharma.flow
    dIpharma <- SIpharma.flow  #risk*lambda*Sph
    dSclinic <- -SIclinic.flow
    dIclinic <- SIclinic.flow  #lambda*Spl

    #Output
    list(c(dSpharma,dIpharma,
           dSclinic,dIclinic,
           SIpharma.flow,SIclinic.flow))
  })
}
```

### Data manipulation  

This function just takes the model output (a `mod` file) and uses the data to create other data (e.g. incidence) for plotting.

```{r data}

mod.manipulate <- function(mod){
  
#Instantaneous incidence (hazard) estimates, per 100 person years
#Instantaneous incidence / hazard
mod <- mutate_epi(mod, rate.Pharma = (SIpharma.flow/Spharma)*365*100)
mod <- mutate_epi(mod, rate.Clinic = (SIclinic.flow/Sclinic)*365*100)

#Cumulative incidence
mod <- mutate_epi(mod, cumul.Spharma = cumsum(Spharma))
mod <- mutate_epi(mod, cumul.Sclinic = cumsum(Sclinic))
mod <- mutate_epi(mod, cumul.rate.Pharma = (Ipharma/cumul.Spharma)*365*100)
mod <- mutate_epi(mod, cumul.rate.Clinic = (Iclinic/cumul.Sclinic)*365*100)

return(mod)
}
```

### Initial parameter settings

```{r parameters}

beta <- 0.001   #transmission rate (per contact)
c <- 90/365    #contact rate (contacts per day)
prev <- 0.05    #needs some consideration
#prev <- 0.01
lambda <- beta*c*prev
#lambda <- 0.000008398975
#epsilon <- 0.0  #0 per contact vaccine efficacy in this model usage
risk <- 2.0   #risk multiplier
```

### Running the model  

We are using the EpiModel framework, http://www.epimodel.org/, with help from Sam Jenness (Emory University).  

Note that this code chunk includes the model initiation piece, in which among other things we initialize the overall population size and the size of any sub-populations defined by exposure risk. 
  
```{r EpiModel_deterministic}

param <- param.dcm(lambda = lambda, risk = risk)
init <- init.dcm(Spharma = 10000, Ipharma = 0, #high risk (pharmacies)
                 Sclinic = 10000, Iclinic = 0, #low risk (clinics)
                 SIpharma.flow = 0, SIclinic.flow = 0)
control <- control.dcm(nsteps = 365*1, new.mod = si_ode)
mod <- dcm(param, init, control)
mod

mod <- mod.manipulate(mod)

```

### Model plots  

First let's show the cumulative infections over the course of a year. The population size is 10000. 

The clinic incidence is qualitatively calibrated with this first pass. The pharmacy incidence is modified with the risk multiplier (and not specifying whether this risk is elevated due to individual contact rate, the per-contact transmission probability, or the prevalence of viremia in the contacts of the high risk individuals).  
 
```{r}

plot(mod, y = c("Ipharma", "Iclinic"), 
     alpha = 0.8, 
     main = "Cumulative infections",
     legend = FALSE,
     ylab = "infected",
     xlab = "days",
     col = c("blue", "red"))
legend("bottomright", legend = c("pharmacies", "clinics"), col = c("blue", "red"), lwd = 2, cex = 0.9, bty = "n")
```

Let's just plot hazard/instantaneous incidence in these two populations.  

```{r}

#mod <- mod.manipulate(mod)

plot(mod, y=c("rate.Pharma", "rate.Clinic"),
     alpha = 0.8,
     ylim = c(0, 5),
     main = "Instantaneous incidence",
     xlab = "days",
     ylab = "infections per 100 person yrs",
     legend = FALSE,
     col = c("blue", "red"))
legend("topright", legend = c("Pharmacies", "Clinics"), col = c("blue", "red"), lwd = 2, cex = 0.9, bty = "n")

par(mfrow=c(1,1))
```

Let's just plot cumulative incidence in these two populations.  

```{r}

#mod <- mod.manipulate(mod)

plot(mod, y=c("cumul.rate.Pharma", "cumul.rate.Clinic"),
     alpha = 0.8,
     ylim = c(0, 5),
     main = "Cumulative incidence",
     xlab = "days",
     ylab = "infections per 100 person yrs",
     legend = FALSE,
     col = c("blue", "red"))
legend("topright", legend = c("Pharmacies", "Clinics"), col = c("blue", "red"), lwd = 2, cex = 0.9, bty = "n")

par(mfrow=c(1,1))
```
